---
title: "IE 360- HW 1"
author: "Tugce Caglayan"
date: '2022-04-14'
output: html_document
---

# INTRODUCTION

Economic data can be analyzed as time series data after a certain time and used in forecasting applications. Observed results can depend on many different variables. There may be relationships between variables such as cause-effect or correlation. In this way, more accurate predictions are tried to be obtained for forecasting. In this analysis, the relationship between four different variables-exchange rates (USD/TRY Selling),consumer price index for food,interest rate for personal bank loan,Google search for 'Market Prices'- and their changes over time are examined. In this study, the following questions will be examined:

1-How has the USD/TRY rate,interest rate and cpi changed from January 2020 to today?
2-How can the relationship between Interest Rate and USD/TRY be explained?
3-How was CPI affected by interest rate and USD/TRY change?
4-How is the relationship between these variables in terms of correlation?

In addition, the 'Market Prices' change in Google will be examined together with these data sets.
All data sets are obtained from Electronic Data Delivery System of Central Bank of the Republic of Turkey. 


```{r warning=FALSE, message=FALSE}
library(openxlsx)
library(ggplot2)
library(data.table)
library(skimr)
library(GGally)
library(ggcorrplot)
library(lubridate)
library(readxl)
library(xts)

Exchange_path='C:/Users/tugce/Desktop/SPRING 2022/IE 360/HW 1/ExchangeRate.xlsx'
Interest_path='C:/Users/tugce/Desktop/SPRING 2022/IE 360/HW 1/Interest Rates.xlsx'
CPI_path='C:/Users/tugce/Desktop/SPRING 2022/IE 360/HW 1/ConsumerPriceIndex.xlsx'
market_fiyatlari_path='C:/Users/tugce/Desktop/SPRING 2022/IE 360/HW 1/multiTimeline.xlsx'
Exchange_Rate=read_excel(Exchange_path)
Interest_Rate=read_excel(Interest_path)
CPI=read_excel(CPI_path)
Market_fiyatlari=read_excel(market_fiyatlari_path)

CPI$Date=as.yearmon(CPI$Date)
Exchange_Rate$Date=as.yearmon(Exchange_Rate$Date)
Interest_Rate$Date=as.yearmon(Interest_Rate$Date)
Market_fiyatlari$Date=as.yearmon(Market_fiyatlari$Date)

CPI=data.table(CPI)
Exchange_Rate=data.table(Exchange_Rate)
Interest_Rate=data.table(Interest_Rate)
Market_fiyatlari=data.table(Market_fiyatlari)

```
# DATA VISUALIZATION AND INTERPRETATION 

## 1. Analysis of USD/TRY Exchange Rate
### 1.1 Time Series Plot of USD/TRY Exchange Rate
```{r cars}
ggplot(Exchange_Rate,aes(x=Date,y=ER,group=1))+ geom_line(color="brown",type="solid",size=1)+ggtitle("USD/TRY Rates Between 2020/01-2022/03")+xlab("Months")+ylab("USD/TRY Exchange Rates")
```
### 1.2 Histogram of USD/TRY Exchange Rate

```{r pressure, echo=FALSE}
ggplot(Exchange_Rate,aes(x=ER))+ geom_histogram(color="pink",fill="pink")+ggtitle(" Histogram of USD/TRY Rates Between 2020/01-2022/03")+xlab("USD/TRY Exchange Rates")
```
### 1.3 Box-Plot of USD/TRY Exchange Rate
```{r}
ggplot(Exchange_Rate,aes(x=ER))+ geom_boxplot(color="brown", fill="brown", alpha=0.5)+ggtitle(" Box-Plot of USD/TRY Rates Between 2020/01-2022/03")+xlab("USD/TRY Exchange Rates")
```

## 2. Analysis of Interest Rate
### 2.1 Time Series Plot of Interest Rate
```{r}
ggplot(Interest_Rate,aes(x=Date,y=IR,group=1))+ geom_line(color="brown",type="solid",size=1)+ggtitle("Personal Bank Loan Interest Rates Between 2020/01-2022/03")+xlab("Years")+ylab("Interest Rates")
```
### 2.2 Histogram of Interest Rate
```{r}
ggplot(Interest_Rate,aes(x=IR))+ geom_histogram(color="pink",fill="pink")+ggtitle(" Histogram of Interest Rates Between 2020/01-2022/03")+xlab("Interest Rates")
```

### 2.3 Box-Plot of Interest Rate
```{r}
ggplot(Interest_Rate,aes(x=IR))+ geom_boxplot(color="brown", fill="brown", alpha=0.5)+ggtitle(" Box-Plot of Interest Rates Between 2020/01-2022/03")+xlab("Interest Rates")

```
## 3.Analysis of CPI
### 3.1 Time Series Plot of CPI
```{r}
ggplot(CPI,aes(x=Date,y=Consumer_Price_Index,group=1))+ geom_line(color="brown",type="solid",size=1)+ggtitle("Consumer Production Index for Food Between 2020/01-2022/03")+xlab("Years")+ylab("CPI")
```
### 3.2 Histogram of CPI
```{r}
ggplot(CPI,aes(x=Consumer_Price_Index))+ geom_histogram(color="pink",fill="pink")+ggtitle(" Histogram of CPI Between 2020/01-2022/03")+xlab("CPI")
```
### 3.3 Box-Plot of CPI
```{r}
ggplot(CPI,aes(x=Consumer_Price_Index))+ geom_boxplot(color="brown", fill="brown", alpha=0.5)+ggtitle("Box-Plot of CPI Between 2020/01-2022/03")+xlab("CPI")
```

## 4.Analysis of 'Market Fiyatları' Search
### 4.1 Time Series Plot of 'Market Fiyatları' Search
```{r}
ggplot(Market_fiyatlari,aes(x=Date,y=MF,group=1))+geom_line(color="brown",type="solid",size=1)+ggtitle(" 'Market Fiyatları' Search on Google Between 2020/01-2022/03")+xlab("Years")+ylab("Market Fiyatları Search")
```
### 4.2 Histogram of 'Market Fiyatları' Search
```{r}
ggplot(Market_fiyatlari,aes(x=MF))+geom_histogram(color="pink",fill="pink")+ggtitle(" Histogram of 'Market Fiyatları' Search Between 2020/01-2022/03")+xlab("'Market Fiyatları' Search")
```

### 4.3 Box-Plot of 'Market Fiyatları' Search
```{r}
ggplot(Market_fiyatlari,aes(x=MF))+geom_boxplot(color="brown", fill="brown", alpha=0.5)+ggtitle(" Box-Plot of 'Market Fiyatları' Search Between 2020/01-2022/03")+xlab("'Market Fiyatları' Search")
```
# CORRELATION ANALYSIS

## Pairwise Correlation Analysis

```{r}
data=merge(Exchange_Rate,Interest_Rate,by='Date')
data_new=merge(data,CPI,by='Date')
data_last=merge(data_new,Market_fiyatlari,by='Date')
data_last
summary(data_last[,c(2:5)])
correlation_analysis=cor(data_last[,c(2:5)])
ggcorrplot(correlation_analysis,hc.order = TRUE,type="lower",lab=TRUE)

```
```{r}
ggpairs(data_last[,c(2:5)])
```
From the correlation figures, it can be seen that CPI for food is highly correlated with the exchange rate. This situation can be explained by the government's import-export policy.The decrease in the food production in our country can explain this. In addition, even if there is a production, the transportation cost required to reach the consumer might be a reason of this relationship since we also import gasoline and vehicles from other countries.
# CONCLUSION


